Evaluating a simplified forecast model in comparison to the ECDC forecasting hub ensemble

### 2022-08-31 paper/paper.Rmd
comments from @bisaloo

These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon. Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available. Typically, the ensemble forecast produced each week comprises at least 5 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].

@@

Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub. 

@@

In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarising its implementation, with a focus on minimal resource use and reproducibility, as a GitHub action. 
We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
We then evaluate its real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
Finally, we summarise our findings, discuss their implications, and highlight areas for potential study. We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models

@@

\begin{align*}
  r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
  r_t &= \left(\mathcal{H}[a](r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
  \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
\end{align*}

@@

Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals [b][c][d][e]and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
This is true across forecast horizons but the magnitude of the different increases. This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecasts we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) as noted elsewhere.
This indicates a right skew in performance for the surrogate model point forecasts not observed for the ensemble forecast which matches that observed when evaluating the full predictive distributions of both forecasts.

@@

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty[f][g].
In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction[h].
Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of forecasts.
The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to increasing incidence rates across forecast locations linked to BA.2.
Unlike our earlier forecasting efforts, the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to overcover at large intervals.
Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of the difference in performance. 
Indeed, the median point performance of the forecasts was comparable though the surrogate model performed substantially worse on average due to its skewed performance distribution.

@@

There are no other studies in the epidemiology literature which we are aware of that attempt to develop a forecasting model based on the observed behaviour of a multi-team, multi-model ensemble.
Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
However, the CDC [@REF], ECDC [@REF], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles. 
In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study) and comparatively little attention has been paid to coverage.
The CDC forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
Similarly, the Germany and Poland, and ECDC forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @REF].
In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions after having concluded it was a favourable choice [@REF]. 
The CDC forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF].
Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases.
That being said no forecast hub has switched to these alternative ensemble designs for their operational forecast implying they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study. 
Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric. 
Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@refNikosinGermany].

@@

[a]Isn't the indicator function with 1(.) just more common and potentially easier to read?
[b]I would be very surprised if that were true. The reason is that the outer intervals get tiny weights in the WIS and performance usually is more strongly influenced by the inner intervals.
[c]This hangs on if the output from score is already weighted or not which I was not clear on?
[d]Which was my understanding of the implementation (weighting happens in score and summarise_score is just an averaging?)

@@

Perhaps there is an implementation nuance? I agree I was surprised by the magnitude of this difference.
[e]Aah I was just looking at Figure 3c here (I think it needs to be 3b). Oooh ok I think I now know what you mean. I would think that Figure 3b) is misleading. For every interval x, this is the score from the surrogate over the score for the ensemble for that interval x, right? Two reasons it's misleading: 
1) I think the values that you are dividing just become very small and therefore what you get is unstable, because the denominator gets smaller and smaller
2) the outer values don't influence the average a lot, because you first add the values for the individual intervals up (to get one WIS for ensemble and surrogate) and then you divide these values. The plot suggests that you divide the values for all the quantiles and then average to get your relative score
[f]WIS decompositions would be nice to look at. Although I guess can be left to future Sam (or never Sam)
[g]Suggest add to slack as a suggestion for discussion there (close once done please).
[h]As Johannes mentioned a while ago: It is a priori unclear what the "correct" distribution of overprediction, underprediction and dispersion is. So I think it is technically not correct to talk about a bias towards over / underprediction. Adding here as a comment because it might be relevant throughout. We could look at the bias value from scoringutils.

>>>>

These forecast collaborations, and other similar infectious disease forecasting efforts, have focussed on forecasting reported cases and deaths by epidemiological week (i.e Saturday to Saturday) over a one to four-week time horizon. Generally, ground truth data is available daily, and is updated in retrospect as new data becomes available. Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@REF].

@@

Our previous methodology also requires significant computational resources running for an hour on a 16-core machine when producing forecasts for the ECDC forecasting hub.[a][b] 

@@

In this study, we define an initial attempt at a surrogate model to replicate the observed behaviour of current multi-team forecast ensembles whilst maintaining a set of clear assumptions and summarising its implementation, with a focus on minimal resource use and reproducibility, as a GitHub Actions workflow. 
We briefly outline some of the summary statistics it produces and their epidemiological relevance. 
We then evaluate its real-time performance in comparison to the ECDC forecast hub ensemble by visualising forecasts, using the weighted interval score [@REF], a commonly used proper scoring rule, and quantifying the empirical coverage of the forecasts produced.
We attempt to highlight settings where this model performs well as a surrogate to the ensemble forecast and areas where more work is needed. 
Finally, we summarise our findings, discuss their implications, and highlight areas for potential study. We aim for this work to highlight some of the potential implicit assumptions of current forecast hub ensembles, provide a sensible, low-resource, surrogate model in settings where large-scale collaborative forecasting efforts are not possible, and provide inspiration for forecasters looking to make principled improvements to their models

@@

\begin{align*}
  r_0 &\sim \text{Normal}\left(0, 0.25 \right) \\
  r_t &= \left(\mathcal{H}[c](r_{t-1} > 0) \xi_{+} + \mathcal{H}(r_{t-1} \le 0) \xi_{-}\right)r_{t-1} + \epsilon_t  \\
  \epsilon_t &= \mathcal{H}(t > 0) \beta \epsilon_{t-1} + \eta_t
\end{align*}

@@

Breaking down the relative weighted interval score by forecast interval we observe that the majority of the difference in overall scores is likely driven by differences in the outer intervals [d][e][f][g][h]and in particular the tails of the forecast (Figure \@ref(fig:plot-coverage) b).
This is true across forecast horizons but the magnitude of the different increases. This implies that the surrogate and ensemble model produced comparably performing point forecasts and indeed if we consider the absolute error of the median forecasts we see that the median for the ensemble forecast was `r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` whilst the median for the surrogate forecast was `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`.
However, when we consider the mean absolute error of the median we find that the ensemble forecast outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`) as noted elsewhere.
This indicates a right skew in performance for the surrogate model point forecasts not observed for the ensemble forecast which matches that observed when evaluating the full predictive distributions of both forecasts.

@@

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble though with greater uncertainty.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction[i].
Evaluating the relative performance of the surrogate model compared to the ECDC forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of forecasts.
The median relative performance was only marginally worse across forecast horizons with the distribution of relative scores having a heavy right tail.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to increasing incidence rates across forecast locations linked to BA.2.
Unlike our earlier forecasting efforts, the effect of outlier forecasts was much reduced though the surrogate model was less likely to outperform the ensemble than our earlier forecasting model methodology [@refNikosinGermany].
In general, the relative performance of the surrogate model was stable across forecast horizons, unlike our previous methods which rapidly became less performant, though the distribution of relative performance did have an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger credible intervals whereas the surrogate model was substantially better calibrated for all targets though it tended to overcover at large intervals.
Stratifying the relative weighted interval score by interval we found that larger intervals (i.e the tails of the forecast distribution) were responsible for the majority of the difference in performance. 
Indeed, the median point performance of the forecasts was comparable though the surrogate model performed substantially worse on average due to its skewed performance distribution.

@@

There are no other studies in the epidemiology literature which we are aware of that attempt to develop a forecasting model based on the observed behaviour of a multi-team, multi-model ensemble.
Few studies focus on delivering computationally feasible forecasting models in a reproducible framework backed by an openly accessible compute platform. 
However, the CDC [@REF], ECDC [@REF], and Germany/Poland [@REF] forecasting hubs have published a range of evaluations of forecasts submitted to their platforms and the relative performance of their ensembles. 
In general, these studies have struggled to draw general conclusions about the structural assumptions of forecast models they consider "good" (generally they have defined this as minimising the weighted interval score as done in this study) and comparatively little attention has been paid to coverage.
The CDC forecast hub identified the top 5 performing models and noted the structural assumptions they made [@REF].
However, they did not extensively compare and contrast these conclusions to arrive at a set of desired forecast assumptions as done in this study to motivate the surrogate model or explore the performance of a forecasting model designed with these assumptions in mind.
Similarly, the Germany and Poland, and ECDC forecasting hubs were able to identify forecast models that performed comparably as well as their ensemble forecasts but did not derive structural assumptions that led to this out-performance or detail explicitly what the desirable performance characteristics would be, aside from optimising the weighted interval score.
All comparable forecast hub projects found that their ensemble was often the best choice, had desirable characteristics such as robustness, though this was rarely fully defined, and should be the output used by forecast consumers [@REF; @REF; @REF].
In general, during the study period, all projects used the same unweighted median ensemble forecast of all submissions after having concluded it was a favourable choice [@REF]. 
The CDC forecasting hub also evaluated a range of other ensemble approaches, such as inverse weighted interval score weighting, unweighted ensembles of a selection of models based on recent performance, and mean ensembling [@REF].
Work on this is still ongoing but these more complex ensembling approaches were shown to outperform the median of all submitted forecasts in many cases.
That being said no forecast hub has switched to these alternative ensemble designs for their operational forecast suggesting they do not think the evidence base is strong enough for them to be used by forecast consumers and hence the median of all submitted forecasts remains the community-suggested default ensemble option and a sensible target for our study. 
Other studies have been published evaluating single forecast models in comparison to ensemble performance from the forecast hub but in general, these have not focussed on replicating ensemble behaviour but rather optimising the target evaluation metric. 
Our previous work also highlighted the lack of calibration in an ensemble forecast from the Germany/Poland forecasting hub compared to forecasts from epidemiological models and noted the bias towards underprediction observed in the ensemble forecasts and not the model-based forecasts [@refNikosinGermany].

@@

[a]For this to be really informative, you would need to specify the CPU clock speed. There are huge disparities between the most recent 16-core machines and older ones.
[b]Agree it could/should list the Azure type and clock speed but disagree it brings no information as the comparison is to a 2 core CI process available in 2022 (i.e when the paper is written).

I think perhaps adding this detail could be a bit overwhelming for an introduction but worth adding to try it out
[c]Isn't the indicator function with 1(.) just more common and potentially easier to read?
[d]I would be very surprised if that were true. The reason is that the outer intervals get tiny weights in the WIS and performance usually is more strongly influenced by the inner intervals.
[e]This hangs on if the output from score is already weighted or not which I was not clear on?
[f]Which was my understanding of the implementation (weighting happens in score and summarise_score is just an averaging?)

@@

Perhaps there is an implementation nuance? I agree I was surprised by the magnitude of this difference.
[g]Aah I was just looking at Figure 3c here (I think it needs to be 3b). Oooh ok I think I now know what you mean. I would think that Figure 3b) is misleading. For every interval x, this is the score from the surrogate over the score for the ensemble for that interval x, right? Two reasons it's misleading: 
1) I think the values that you are dividing just become very small and therefore what you get is unstable, because the denominator gets smaller and smaller
2) the outer values don't influence the average a lot, because you first add the values for the individual intervals up (to get one WIS for ensemble and surrogate) and then you divide these values. The plot suggests that you divide the values for all the quantiles and then average to get your relative score
[h]That sounds like a wording/interpretation issue vs a change the plot issue?

Do you have suggestions for that wording or if suggesting the plot is changed what it should be changed to?
[i]As Johannes mentioned a while ago: It is a priori unclear what the "correct" distribution of overprediction, underprediction and dispersion is. So I think it is technically not correct to talk about a bias towards over / underprediction. Adding here as a comment because it might be relevant throughout. We could look at the bias value from scoringutils.
